---
title: "run_analysis.Rmd"
date: "Monday, April 18, 2016"
output: html_document
---

**STEP #1 LOAD PACKAGE** 

```{r}
packages <- c("data.table", "reshape2")
sapply(packages, require, character.only = TRUE, quietly = TRUE)
```

**STEP #2 GET PATH**

```{r, echo=TRUE}
path = getwd()
```

**STEP #5a READ SUBJECTS**

```{r, echo=TRUE}
pathIn <- file.path(path, "UCI HAR Dataset")
list.files(pathIn, recursive = TRUE)
dtSubjectTrain <- fread(file.path(pathIn, "train", "subject_train.txt"))
dtSubjectTest <- fread(file.path(pathIn, "test", "subject_test.txt"))
```

**STEP #5b READ ACTIVITIES**

```{r, echo=TRUE}
dtActivityTrain <- fread(file.path(pathIn, "train", "Y_train.txt"))
dtActivityTest <- fread(file.path(pathIn, "test", "Y_test.txt"))
```

**STEP #6 READ DATA FILES**

```{r, echo=TRUE}
fileToDataTable <- function(f) {
    df <- read.table(f)
    dt <- data.table(df)
}
dtTrain <- fileToDataTable(file.path(pathIn, "train", "X_train.txt"))
dtTest <- fileToDataTable(file.path(pathIn, "test", "X_test.txt"))
```

**STEP #7 MERGE TRAINING AND TEST**

```{r, echo=TRUE}
dtSubject <- rbind(dtSubjectTrain, dtSubjectTest)
setnames(dtSubject, "V1", "subject")
dtActivity <- rbind(dtActivityTrain, dtActivityTest)
setnames(dtActivity, "V1", "activityNum")
dt <- rbind(dtTrain, dtTest)
```

**STEP #8 MERGE COLUMNS, SET KEY**

```{r, echo=TRUE}

dtSubject <- cbind(dtSubject, dtActivity)
dt <- cbind(dtSubject, dt)

setkey(dt, subject, activityNum) # sort by key

```

**STEP #9 GET FEATURE NAMES**

```{r, echo=TRUE}
dtFeatures <- fread(file.path(pathIn, "features.txt"))
setnames(dtFeatures, names(dtFeatures), c("featureNum", "featureName")) #renames names(dtFeatures) with the new char vector names
```

**STEP #10 GET MEASUREMENTS FOR MEAN AND STD**

```{r, echo=TRUE}
dtFeatures <- dtFeatures[grepl("mean\\(\\)|std\\(\\)", featureName)]
```

**STEP #11 CREATE FEATURE CODE**

```{r, echo=TRUE}
dtFeatures$featureCode <- dtFeatures[, paste0("V", featureNum)]
head(dtFeatures)
```

**STEP #12 GET MEASUREMENTS FOR MEAN AND STD**

```{r, echo=TRUE}
select <- c(key(dt), dtFeatures$featureCode)
dt <- dt[, select, with = FALSE]
```

**STEP #13 CHANGE TO DESCRIPTIVE NAMES**

```{r, echo=TRUE}
dtActivityNames <- fread(file.path(pathIn, "activity_labels.txt"))
setnames(dtActivityNames, names(dtActivityNames), c("activityNum", "activityName"))
```

**STEP #14 MERGE ACTIVITY LABELS**

```{r, echo=TRUE}
dt <- merge(dt, dtActivityNames, by = "activityNum", all.x = TRUE)
```

**STEP #15 SET ACTIVITYNAME AS KEY**

```{r, echo=TRUE}
setkey(dt, subject, activityNum, activityName)
```

**STEP #16 MELT DATA FROM SHORT AND WIDE FORMAT TO TALL AND NARROW FORMAT**

```{r, echo=TRUE}
dt <- data.table(melt(dt, key(dt), variable.name = "featureCode"))
```

**STEP #17 MERGE ACTIVITY NAME**

```{r, echo=TRUE}
dt <- merge(dt, dtFeatures[, list(featureNum, featureCode, featureName)], by = "featureCode", all.x = TRUE)
```

**STEP #18 CREATE FACTOR VARIABLES**

```{r, echo=TRUE}
dt$activity <- factor(dt$activityName)
dt$feature <- factor(dt$featureName)
```

**STEP #19 SEPARATE FEATURES FROM FEATURE NAME**

```{r, echo=TRUE}
dtTidy = unique(dt[,.(activity,subject,featureName,mean = mean(value))])

```


**STEP #22 SAVE DATA**

```{r, echo=TRUE}
f <- file.path(path, "HumanActivitiesFromSmartphoneData.txt")
write.table(dtTidy, f, quote=FALSE, sep="\t", row.names=FALSE)
```


